---
output:
  md_document:
    variant: gfm
html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

#  ggmsa: Plot multiple sequence alignment using ggplot2

[![CRAN_Release_Badge](https://www.r-pkg.org/badges/version-ago/ggmsa)](https://cran.r-project.org/package=ggmsa)
[![CRAN_Download_Badge](https://cranlogs.r-pkg.org/badges/grand-total/ggmsa?color=green)](https://cran.r-project.org/package=ggmsa)


**ggmsa** supports visualizing multiple sequence alignment of DNA and protein sequences using ggplot2. It supports a number of colour schemes, including Chemistry, Clustal, Shapely, Taylor and Zappo.

##  Quick Example 
Plot multiple sequence alignment(colour scheme = 'Chemistry').

###  protein sequences
```{r fig.height = 2, fig.width = 10, message=FALSE, warning=FALSE}
library(ggmsa)
library(ggplot2)
protein_sequences <- system.file("extdata", "sample.fasta", package = "ggmsa")
ggmsa(protein_sequences, 164, 213, color = "Chemistry_AA")
```

###  DNA sequences
```{r fig.height = 6, fig.width = 13, message=FALSE, warning=FALSE}
nt_sequence <- system.file("extdata", "LeaderRepeat_All.fa", package = "ggmsa")
ggmsa(nt_sequence, color = "Chemistry_NT")
```

##  Visualizing multiple sequence alignment with sequence logo.
```{r fig.height=6, fig.width=13, message=FALSE, warning=FALSE}
f <- system.file("extdata", "LeaderRepeat_All.fa", package = "ggmsa")
ggmsa(f, font = NULL, color = "Chemistry_NT" ) + geom_seqlogo()
```

##  Visualizing multiple sequence alignment with GC content.
```{r fig.height=6, fig.width=13, message=FALSE, warning=FALSE}
f <- system.file("extdata", "LeaderRepeat_All.fa", package = "ggmsa")
ggmsa(f, font = NULL, color = "Chemistry_NT" ) + geom_GC()
```

##  Visualizing multiple sequence alignment with ggtree.
```{r fig.height=2, fig.width=10, message=FALSE, warning=FALSE}
library(Biostrings)
x <- readAAStringSet(protein_sequences)
d <- as.dist(stringDist(x, method = "hamming")/width(x)[1])
library(ape)
tree <- bionj(d)
library(ggtree)
p <- ggtree(tree ) + geom_tiplab()

data = tidy_msa(x, 164, 213)
p + geom_facet(geom = geom_msa, data = data,  panel = 'msa',
               font = NULL, color = "Chemistry_AA") +
    xlim_tree(1)
```

## Learn more

For more details, please refer to the [online vignette](https://cran.r-project.org/web/packages/ggmsa/vignettes/ggmsa.html)

If you have installed it, you can also view the vignette on local.

```r
browseVignettes("ggmsa")
```
