---
title: "Theme"
output: 
  prettydoc::html_pretty:
    toc: true
    theme: hpstr
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{ggmsa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

 protein_sequences <- system.file("extdata", "sample.fasta", package = "ggmsa")
 miRNA_sequences <- system.file("extdata", "seedSample.fa", package = "ggmsa")
 nt_sequences <- system.file("extdata", "LeaderRepeat_All.fa", package = "ggmsa")

library(ggmsa)
library(ggplot2)
```

##  Modify components of MSA theme

In this package, there are various powerful parameters to customize the non-data components of MSA plots: i.e. font size, background, sequence names, legends and border. Some parameters can be used to give MSA plots a consistent customized look.   

###  Characters width

**Characters width** can be specified by `char_width`. Default is 0.9.
```{r fig.height = 3, fig.width = 10, warning=FALSE}
ggmsa(protein_sequences, start = 320, end = 360, char_width = 0.5)
```

###  Background 

If `none_bg = TRUE`, only show the character but not the colored **background.**.
```{r fig.height = 3, fig.width = 10, warning=FALSE}
ggmsa(protein_sequences, start = 320, end = 360, none_bg = TRUE) + theme_void()
```

###  Sequence names

 If `seq_name = TRUE`, **the sequence name** will be displayed.  
```{r fig.height = 3, fig.width = 10, warning=FALSE}
ggmsa(protein_sequences, 164, 213, seq_name = TRUE)
```

### Legend

If `show.legend = TRUE`, the **legend** of MSA plot will be displayed.  
```{r fig.height = 3, fig.width = 10, warning=FALSE}
ggmsa(protein_sequences, 164, 213, font = NULL, show.legend = TRUE)
```

### Border

Change or set colored block **border**. if `border = NA`, the border would be removed.

```{r fig.height = 3, fig.width = 10, warning=FALSE}
ggmsa(protein_sequences, 164, 213, font = NULL, border = NA)
```

Or, modify the border color by assign the color to border, like this： `border = "white"`
```{r fig.height = 3, fig.width = 10, warning=FALSE}
ggmsa(protein_sequences, 164, 213, font = NULL, border = "white")
```

### Position Highlighted

Specified position can be highlighted  but the block colors only assign by ggplot2 automatically. 


```{r fig.height = 3, fig.width = 10, warning=FALSE}
ggmsa(protein_sequences, 164, 213, posHighligthed = c(185, 190), char_width = 0.5)
```



